---
title: "210406_TidyTuesday"
author: "Angela"
date: "April 6, 2021"
output: html_document
---

These are the libraries required:

```{r}
library(tidytuesdayR)
library(treemap)
library(tidyverse)
library(gifski)
library(glue)
library(RColorBrewer)

```


The data can be loaded in via the TidyTuesday library.

```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 15)
```


As a first step rename the columns to make them more intuitive to work with.

```{r}
df_brazil_loss <- data.frame(tuesdata[3])
df_brazil_loss <- df_brazil_loss %>%
  rename("Year" = brazil_loss.year, 
         "Commercial crops" = brazil_loss.commercial_crops, 
         "Flooding due to dams" = brazil_loss.flooding_due_to_dams, 
         "Natural disturbances" = brazil_loss.natural_disturbances, 
         "Pasture" = brazil_loss.pasture,
         "Selective logging" = brazil_loss.selective_logging,
         "Fire" = brazil_loss.fire,
         "Mining" = brazil_loss.mining,
         "Other infrastructure" = brazil_loss.other_infrastructure,
         "Roads" = brazil_loss.roads,
         "Tree plantations" = brazil_loss.tree_plantations_including_palm,
         "Small scale clearing" = brazil_loss.small_scale_clearing) %>%
head(df_brazil_loss)
```


Then change the format of the data frame and add in additional columns as required for the plots.

```{r}
df_brazil_loss_long <- df_brazil_loss %>% 
  select(c(3:14)) %>%
  pivot_longer(!Year, names_to = "Drivers of loss", values_to = "Area in hectars") %>%
  group_by(Year) %>%
  mutate("Annual Total" = sum(`Area in hectars`)) %>%
  ungroup() %>%
  mutate("Percentage" = round(`Area in hectars`/`Annual Total`*100, 2),
         "Drivers of loss" = paste(`Drivers of loss`, glue("{`Percentage`}%")))

df_brazil_loss_long
```
    

Here's what one treemap looks like.

```{r}
  df_brazil_loss_long %>%
  filter(Year == 2013) %>% 
  treemap(index = "Drivers of loss",
          vSize = "Area in hectars",
          type = "index",
          palette = "RdGy",
          title = "Drivers for Area of Rainforest lost in Brazil in 2013",
          algorithm = "pivotSize",
          sortID = "Drivers of loss")
```


This loop creates one treemap for each year and saves it in a folder called 'pngs'.

```{r}
png_files <- character()
for (i in unique(df_brazil_loss_long$Year)) {
  png(glue("pngs/treemap{i}.png"), width = 750, height = 500, res = 140)
  df_brazil_loss_long %>%
  filter(Year == i) %>% 
  treemap(index = "Drivers of loss",
          vSize = "Area in hectars",
          type = "index",
          palette = "RdGy",
          title = glue("Drivers for Area of Rainforest lost in Brazil in {i}"),
          algorithm = "pivotSize",
          sortID = "Drivers of loss")
  dev.off()
  png_path <- glue("pngs/treemap{i}.png")
  png_files[i-2000] <- png_path
}
```


With this code, the pngs are transformed into one gif file, which is saved in the same folder where the Markdown document is.

```{r}
gifski(
png_files,
gif_file = "drivers_rainforest_loss.gif",
width = 800,
height = 600,
delay = 1.5,
loop = TRUE,
progress = FALSE
)

```

